---
title: "LiveAssignment4"
author: "William Arnost"
date: "June 3, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Harry Potter Cast (50%)  
There is a lot of text associated with this question. We are scraping the cast list for Harry Potter 2 and cleaning the data. Here is a summary of the requrirements:  
1. Scrape the data from the provided link and put it into a data table  
2. Clean the table  
    a. It should not have blank or otherwise useless data  
    b. It should have intuitive column names (ideally 2 to start - Actor and Character)   
    c. Fix the issue with Mr. Warwick's double character role  
    d. remove the "Rest of cast..." observation that doesn't belong  
3. Split the actor names into first name / surname  
4. Display the first 10 rows of the cleaned data, 3 relevant columns only.  

Here are the packages and URL I will be using
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
library(humaniformat)
library(stringi)

url<-"https://www.imdb.com/title/tt1201607/fullcredits?ref_=tt_ql_1"
```

Here I am extracting the table from the html, checking the values, and doing some cleaning
```{r}
## Read the html, extract the table nodes, then the specific table of interest
cast_list <- read_html(url) %>% html_nodes(.,"table") %>% html_table %>%.[[3]]

##Take a peek at the data
str(cast_list)

## Make sure X1 and X3 have no meaningful data
table(cast_list$X1)
table(cast_list$X3)

## Get rid of X1 and X3
cast_list$X1 <- NULL
cast_list$X3 <- NULL

## Assign meaningful names
names(cast_list) <- c("Actor","Character")

##Remove junk rows and remove extra spaces and \n from Character field
cast_list_filtered <- cast_list %>% filter(Actor != "",Actor !="Rest of cast listed alphabetically:")
cast_list_filtered$Character <- str_squish(cast_list_filtered$Character)
str(cast_list_filtered)
head(cast_list_filtered,10)
```

Names are super tricky. I am going to use a parse function from the humaniformat package. It doesn't handle accents well so I need to remove those, I hope no one is offended.
```{r}
## Removing Accents, then parsing names
no_accents <- stri_trans_general(cast_list_filtered$Actor,"Latin-ASCII")
parsed_names <- parse_names(no_accents)

## If there is a middle name, combine with the first name. 
firstName <- ifelse(is.na(parsed_names$middle_name),
                    parsed_names$first_name,
                    paste(parsed_names$first_name,parsed_names$middle_name)
                    )
lastName <- parsed_names$last_name

## Add the clean names to the previous data frame
cast_list_addNames <- cast_list_filtered %>% add_column(FirstName = firstName, Surname = lastName)
head(cast_list_addNames)
```

Homenum revelio. Here is my cleaned data.
```{r}
cast_list_final <- cast_list_addNames %>% select(FirstName, Surname, Character)
head(cast_list_final,10)
```


	
#	2. SportsBall (50%)  

There is a lot of text associated with this question. We are scraping the shooting statistics for the San Antonio Spurs and cleaning the data. Here is a summary of the requrirements:  
1. Scrape the shooting statistics table for the San Antonio Spurs  
2. Clean the table so it shows one row per player. Make sure to avoid blank data, extra column names, and get rid of the total row.  
3. Split the player name into name and position  
4. verify numeric columns are numeric  
5. Create a colorful bar chart that shows the Field Goals Percentage Per Game for each person.   
    a. Informative Title, centered  
    b. Relevant x and y axis labels (not simply variables names!)  
    c. Human-readable axes with no overlap  
    d. Color the columns by the team member's position  
  
First step, get the html data, parse our the tables and get them into data frames.
```{r}
url <- "http://www.espn.com/nba/team/stats/_/name/sa/san-antonio-spurs"

## Read the html, get the table nodes, put the data into tables.
html_tables <- read_html(url) %>% html_nodes(.,"table") %>% html_table(., fill = TRUE)

## Extract the desired info
stats <- data.frame(html_tables[8])
players <- data.frame(html_tables[6])
```

Now lets combine the Stats and Players table, then separate name and position
```{r}
names(players) <- "Player"
shooting_stats <- cbind(players,stats) %>% filter(Player != "Total")

## This removes the last word from the name, getting rid of position
shooting_stats$Name = word(shooting_stats$Player,1,-2)

## This grabs the last word, which is the position
shooting_stats$Position = word(shooting_stats$Player,-1,-1)

## Reordering the columns
shooting_stats_final <- shooting_stats %>% select(Name,Position,2:14)
str(shooting_stats_final)
```

Notice in the previous table that str shows us the numeric columns are indeed numeric. Now lets make our colorful graph. I have rearranged the names to be in descending order by field goal percentage. All positions are represented in the top 5 so it seems like you can score from any position, at least for the Spurs.
```{r}
ggplot(shooting_stats_final,aes(x=reorder(Name,FG.), y=FG., fill=Position)) + 
  geom_bar(stat = "identity") + 
  coord_flip() + 
  ggtitle("Field Goals Percentage Per Game for the San Antonio Spurs") + 
  ylab("Field Goal Percentage Per Game") + 
  xlab("Player Name") +
  theme(plot.title = element_text(hjust = 0.5),axis.title.y=element_text(angle=0,vjust=1)) 
```


  
  