---
title: "CaseStudy"
author: "William Arnost"
date: "6/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## 0. EDA
```{r}
setwd("C:/Users/William/OneDrive/MSDS_6306_DoingDataScience/MSDS_6306_Homework/CaseStudy01")
library(tidyverse)
library(DataExplorer)

beers <- read_csv("Beers.csv")
breweries <- read_csv("Breweries.csv")
```

```{r}
plot_missing(beers)
plot_missing(breweries)
```

```{r}
beers_unique_val_count <- sapply(beers, function(x) length(unique(x)))
beers_unique_val_count
brew_unique_val_count <- sapply(breweries, function(x) length(unique(x)))
brew_unique_val_count
beers_missing_count <- sapply(beers, function(x) sum(is.na(x)))
beers_missing_count
brew_missing_count <- sapply(breweries, function(x) sum(is.na(x)))
brew_missing_count
```

```{r}
ggplot(beers, aes(x=ABV)) + geom_histogram()
ggplot(beers, aes(x=IBU)) + geom_histogram()
ggplot(beers, aes(as.character(Ounces))) + geom_bar()
ggplot(beers, aes(Style)) + geom_bar() + coord_flip()
ggplot(breweries, aes(State)) + geom_bar() + coord_flip()
table(breweries$State)
```

```{r}
combined <- beers %>% left_join(breweries, by = c("Brewery_id" = "Brew_ID"))
byState <- combined %>% select(State, Brewery_id, ABV, IBU) %>% group_by(State) %>% 
  summarise(Brewery_Count = n(), Median_ABV = median(ABV,na.rm=TRUE), Median_IBU = median(IBU,na.rm=TRUE), Max_ABV = max(ABV,na.rm=TRUE), Max_IBU = max(IBU,na.rm=TRUE))
```
## 1. How many breweries are present in each state?
```{r}
byState %>% select(State, Brewery_Count)
```



## 2. Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.
```{r}
## We did the merge in the EDA step
head(combined,6)
tail(combined,6)
```

## 3.   Report the number of NA's in each column.
```{r}
sapply(combined, function(x) sum(is.na(x)))
```

## 4.   Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
```{r}
ggplot(byState,aes(x=reorder(State,Median_ABV), y=Median_ABV)) + geom_bar(stat="identity") + coord_flip()
ggplot(byState,aes(x=reorder(State,Median_IBU), y=Median_IBU)) + geom_bar(stat="identity") + coord_flip()
```

## 5.   Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r}
byState[which.max(byState$Max_ABV),]
byState[which.max(byState$Max_IBU),]
```

## 6.   Summary statistics for the ABV variable.
```{r}
summary(combined$ABV)
```

## 7.   Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.
```{r}
ggplot(combined,aes(x=IBU,y=ABV)) + geom_point() + geom_smooth(method=lm) 
```
