---
title: "Quandl API"
author: "William Arnost"
date: "May 29, 2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Quandl

The Quandl API allows access to financial data from many different sources. You can check out their website here: https://www.quandl.com/tools/r

##GDP Data using Quandl
Lets pull in US GDP Data using the API. This is sourced from the Federal Reserve.
```{r, message=FALSE,warning=FALSE}
library(Quandl)
data <- Quandl("FRED/GDP", type="raw")
str(data)
head(data)
```

##Transformations
We can also apply transformations to the data, in this case lets just look at the absolute difference between quarters, or how much GDP increased/decreased from the prior period. We can also filter the time period, collapse time periods, and specify other transformations. 
```{r}
data_diff <- Quandl("FRED/GDP", transform="diff")
head(data_diff)
```

##Plot GDP
For fun lets look at a plot of GDP.
```{r,echo=FALSE}
library(ggplot2)

ggplot(data, aes(x=Date, y=Value)) + geom_line() + 
  labs(title="US GDP 1947-Present, Billions of Dollars", x="Year", y="")
```

